Topic modeling on EU and the USA research projects

### 2017-06-27 thesis_report/draft_thesis.tex
Fix 5.1

In this work, we discover topics and their evolution from 1994 to
2017, using the projects' abstracts that got funded in this
period. Using a topic model, we extract a set of topics and
investigate the changes in the content per year/framework programme.
These topics are used to gain insight into some of the research that
occurs the last two decades. Our goal is to share these insights with
researchers and encourage them to develop better strategies to
establish themselves as more effective researchers globally, by
choosing a subject that is more likely to get a funding.

@@

In figures \ref{wordcloudEU4} - \ref{wordcloudEU20} we show the 10 topics from FP4 to H2020 of the EU dataset, described by
the 20 top words with the highest posterior probability of belonging
to the topic. The size of each word is in proportion to each
probability; the bigger the word, the higher the probability. 
Each wordcloud describes a topic, which we can find intuitively (the label 
of each wordcloud corresponds to the title of the topic).

@@

The topics that we are going to discuss in this section are the ones that
stand out in most of the Framework Programmes.
In general, the most popular subject in European Union over the years
is \emph{Information systems}. In our dataset we found projects
related to almost all kinds of information systems, and especially 
\emph{Management information
systems}, \emph{Network information systems} and \emph{Information systems as a
service} to all kinds of organizations, including universities and
companies. Since technology has an impact in every aspect nowadays, it
is normal to have such popularity over the years and it's becoming
popular in education, as well. The same period of time, \emph{Innovation in
Business} emerged as a trending topic, with projects related to new
technologies and software systems in the industry and how they assist
the businesses and the market. In addition, until 2002 a very popular
subject related to \emph{Information systems}, was \emph{Applications in mobile
networks}, \emph{embedded and broadband systems}, which nowadays has lost part
of its popularity in research.

Another popular subject that appears the last 22 years is \emph{Materials research};
the projects related to material research are mostly related to the discovery
and design of new materials, with an emphasis on solids. Materials properties
and performance are used in many disciplines, like chemistry, physics and
engineering.

The third most popular subject is \emph{Molecular biology \& Genetics}.
These projects are related to the cellular molecules that carry out
the biological processes essential for the cell's functions and
maintenance. Especially from 1998 up to day, there is a focus in
Health care and Diagnostics, and for the last decade researchers are
studying the \emph{Cancer detection \& treatment}, which is a common cause
of death in EU and the second most common cause of death in USA.

Moreover, \emph{Environmental, Social and Economic Sustainability} is a
subject very important to the research community and many researchers
work on this up to day. 
The European Union is considered by some to have the most extensive
environmental laws of any international organization, started by the
European Economic Community (EEC), which is the beginning of the EU's
environmental policy. \emph{Social and Economic policies} are also popular in
the EU region over the last decade. Also, the last two decades there
is a focus on the \emph{Green energy production}, like hydropower, solar and
nuclear energy.

During the 6th and the 7th framework programme, \emph{International Research \& 
conferences} came up in our dataset as a trending topic. Researchers have a great
interest to collaborate with other colleagues from all over the world and share
their knowledge. A big part of the projects funded by the EU that is related to
\emph{international research}, is also related to \emph{Physics}, in general, 
and \emph{High energy physics}.

@@

In figures \ref{wordcloudUSA4} - \ref{wordcloudUSA20} we show the 10 topics 
from 1994 to 2017 of the USA dataset, described by
the 20 top words with the highest posterior probability of belonging
to the topic.
By comparing the the areas of interest in the US, through its directorates, 
and our topics, we see that all the topics that we found correspond to either 
a directorate or one of the divisions. Below, we discuss the topics that
stand out in most of the years.

@@

In figure \ref{totaltopicsEU} we show the 10 topics described by
the 20 top words with the highest posterior probability of belonging
to the topic. The size of each word is in proportion to each
probability; the bigger the word, the higher the probability. Each
wordcloud describes a topic, which we can find intuitively. 
We ran the LDA model with 7000
iterations and the same value of the rest of the
parameters, as described in section \ref{modelinference}. Below are
listed the topics that we found corresponding to the words that
describe them:
\begin{itemize}
\item[] Topic 0: Molecular biology
\item[] Topic 1: Materials \& Manufacturing
\item[] Topic 2: Social policy
\item[] Topic 3: Disease detection \& treatment
\item[] Topic 4: Materials science / Physics
\item[] Topic 5: Transport \& Energy
\item[] Topic 6: Business \& Innovation
\item[] Topic 7: Climate change / Environment
\item[] Topic 8: Molecular mechanisms
\item[] Topic 9: Information systems
\item[] Topic 10: International research
\item[] Topic 11: Energy production
\end{itemize}
\begin{center}
\begin{figure}
\begin{tabular}{rcl}
\subfloat[Molecular biology]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic0_all.png}} &
\subfloat[Materials \& Manufacturing]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic1_all.png}} &
\subfloat[Social policy]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic2_all.png}}\\
\subfloat[Disease detection \& treatment]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic3_all.png}} &
\subfloat[Materials science / Physics]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic4_all.png}} &
\subfloat[Transport \& Energy]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic5_all.png}}\\
\subfloat[Business \& Innovation]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic6_all.png}} &
\subfloat[Climate change / Environment]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic7_all.png}} &
\subfloat[Molecular mechanisms]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic8_all.png}}\\
\subfloat[Information systems]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic9_all.png}} &
\subfloat[International research]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic10_all.png}} &
\subfloat[Energy production]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic11_all.png}}\\
\end{tabular}
\caption{Wordclouds of topics of the EU dataset (1994-2017)}
\label{totaltopicsEU}
\end{figure}
\end{center}

>>>>

In this work, we discover topics and their evolution from 1994 to
2017, using the projects' abstracts that got funded in this
period. Using a topic model, we extract a set of topics and
investigate the changes in the content per year/framework programme.
These topics are used to gain insight into some of the research that
occurs the last two decades. We hope to share these insights with
researchers and encourage them to develop better strategies to
establish themselves as more effective researchers globally, by
choosing a subject that is more likely to get a funding.

@@

In figures \ref{wordcloudEU4} - \ref{wordcloudEU20} we show the 10 topics from FP4 to H2020 of the EU dataset, described by
the 20 top words with the highest probability of belonging
to the topic. The size of each word is in proportion to each
probability; the bigger the word, the higher the probability. 
Each wordcloud describes a topic, which we can find intuitively (the label 
of each wordcloud corresponds to the title of the topic).

@@

In order to obtain an overall view of th etopics in EU, we calculated for each
one of the FPs the proportion of the documents-projects that are assigned to the
topics of the FP, normalized by the total number of the documents of the FP. The
topics that we are going to discuss in this section are the ones that either
show up in most of the FPs or they have a high percentage in one of them.

\emph{Molecular biology \& Genetics} is a popular topic that appears in all the
FPs, from 1994 to 2017, where the 13\% of the projects on average belong to this
category. These projects are related to the cellular molecules that carry out
the biological processes essential for the cell's functions and maintenance.
Another popular subject is \emph{Materials research}; the projects related to
material research are mostly related to the discovery and design of new
materials, with an emphasis on solids. Materials properties and performance are
used in many disciplines, like chemistry, physics and engineering.
\emph{Materials} as a topic covers the 13\% of the projects in FP4, while it
doesn't stand out as a distinguish topic in FP5. However, in FP6 the 22\% of the
projects refer to Materials, and since then it gradually loses its popularity,
with 11\% of the projects in FP7 and only 6\% in H2020.

\emph{Management software \& systems} is a subject that appears only in FP4 and
FP5, with the 12\% of the projects. emph{Information systems \& technologies'}
related projects cover the 8,5\% of the projects on average from FP5-H2020.
Since technology has an impact in every aspect nowadays, it is normal to have a
stable popularity over the years.

Moreover, \emph{Environmental Sustainability and policies} is a popular subject
in FP4, with the 12\% of the projects, while it gradually loses its popularity
over the years. The European Union is considered by some to have the most
extensive environmental laws of any international organization, started by the
European Economic Community (EEC), which is the beginning of the EU's
environmental policy. The \emph{Energy production} is a topic that shows up in
FP4 with a 13\% of the projects, while in FP5 the number of the projects related
to \emph{Green energy} (hydropower, solar, nuclear energy) is increasing up to
17,5\% and since then it also loses its populariry. The last 3 years, only the
3\% of the projects are related to \emph{Green energy production}.

In FP6 and FP7, \emph{Social and Economic policies} appears with a small
percentage of the projects (5\%). However, in H2020 it has a big increace, with
the 17\% of the projects related to this topic. The last 3-4 years,
\emph{Business \& Innovation} emerged as a trending topic, with 1 out of 4
projects being related to the support of businesses and the market, with the
assistance of new technologies and innovation.

The last two decades, there is a 10\% of the projects related to \emph{Physics}
and \emph{High energy physics}. Also, \emph{International Research} came up in
our dataset as a trending topic, with a 16\% of projects in FP6 and FP7.
Finally, the last decade, the 5\% of the projects are related to \emph{Cancer
detection \& treatment}, which is a common cause of death in EU and the second
most common cause of death in USA.

@@

In figures \ref{wordcloudUSA4} - \ref{wordcloudUSA20} we show the 10 topics 
from 1994 to 2017 of the USA dataset, described by
the 20 top words with the highest probability.
By comparing the the areas of interest in the US, through its directorates, 
and our topics, we see that all the topics that we found correspond to either 
a directorate or one of the divisions. Below, we discuss the topics that
stand out in most of the years.

@@

In figure \ref{totaltopicsEU} we show the 10 topics described by
the 20 top words with the highest probability of belonging
to the topic. The size of each word is in proportion to each
probability; the bigger the word, the higher the probability. Each
wordcloud describes a topic, which we can find intuitively. 
We ran the LDA model with 7000
iterations and the same value of the rest of the
parameters, as described in section \ref{modelinference}. Below are
listed the topics that we found corresponding to the words that
describe them:
\begin{itemize}
\item[] Topic 0: Molecular biology
\item[] Topic 1: Materials \& Manufacturing
\item[] Topic 2: Social policy
\item[] Topic 3: Disease detection \& treatment
\item[] Topic 4: Materials science / Physics
\item[] Topic 5: Transport \& Energy
\item[] Topic 6: Business \& Innovation
\item[] Topic 7: Climate change / Environment
\item[] Topic 8: Molecular mechanisms
\item[] Topic 9: Information systems
\item[] Topic 10: International research
\item[] Topic 11: Energy production
\end{itemize}
\begin{center}
\begin{figure}
\begin{tabular}{rcl}
\subfloat[Molecular biology]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic0_all.png}} &
\subfloat[Materials \& Manufacturing]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic1_all.png}} &
\subfloat[Social policy]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic2_all.png}}\\
\subfloat[Disease detection \& treatment]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic3_all.png}} &
\subfloat[Materials science / Physics]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic4_all.png}} &
\subfloat[Transport \& Energy]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic5_all.png}}\\
\subfloat[Business \& Innovation]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic6_all.png}} &
\subfloat[Climate change / Environment]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic7_all.png}} &
\subfloat[Molecular mechanisms]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic8_all.png}}\\
\subfloat[Information systems]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic9_all.png}} &
\subfloat[International research]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic10_all.png}} &
\subfloat[Energy production]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic11_all.png}}\\
\end{tabular}
\caption{Wordclouds of topics of the EU dataset (1994-2017)}
\label{totaltopicsEU}
\end{figure}
\end{center}
